Learning Discrete Bayesian Networks from Continuous Data

### 2017-06-06 jair_Learning_Discrete_Bayesian_Networks_from_Continuous_Data.tex
might be last push
last update?
page number and table/figure rearrange

\jairheading{59}{2017}{XXX}{10/16}{05/17}
\ShortHeadings{Learning Discrete Bayesian Networks from Continuous Data}
{Chen, Wheeler, \& Kochenderfer}
\firstpageno{25}

@@

\author{\name Yi-Chun Chen \email yichunc0419@alumni.stanford.edu \\
	\addr Institute of Computational and Mathematical Engineering,\\
	Stanford University, Stanford, CA  94035 USA
	\AND
	\name Tim A. Wheeler \email wheelert@stanford.edu \\
	\addr Department of Aeronautics and Astronautics, Stanford University,\\
	Stanford, CA  94035 USA
	\AND
	\name Mykel J. Kochenderfer \email mykel@stanford.edu \\
	\addr Department of Aeronautics and Astronautics, Stanford University,\\
	Stanford, CA  94035 USA}
%\authorrunning{Short form of author list} % if too long for running head

@@

%%%%%% Old Abstract %%%%%
%Real data often contains a mixture of discrete and continuous variables, but many Bayesian network structure learning and inference algorithms assume all random variables are discrete. Continuous variables are often discretized, but the choice of discretization policy has significant impact on the accuracy, speed, and interpretability of the resulting models. This paper introduces a principled Bayesian discretization method for continuous variables in Bayesian networks with quadratic complexity instead of the cubic complexity of other standard techniques. Empirical demonstrations show that the proposed method is superior to the state of the art. In addition, this paper shows how to incorporate existing methods into the structure learning process to discretize all continuous variables and simultaneously learn Bayesian network structures.
%%%%%%%%%%%%%%%%%%

@@

@@

% Let $n_p$ and $n_c$ respectively refer to the number of parents and number of children of $X$.

% Let $D$ be a dataset of $N$ samples from which the discretization will be learned, sorted in ascending order according to $X$.
% Let $D_{\boldsymbol{Y}}$ will refer to the data instances associated with variables $\boldsymbol{Y}$, thus $D_X = \{ x_1,x_2,x_3,\ldots,x_n \}$.
% In order to facilitate understanding, we temporarily assume all values in $D_X$ are unique, that is, $x_1 < x_2 < \ldots < x_n$.
% It follows that a discretization policy $\Lambda$ on $X$ can be written as $M_X = \brock{n_1,n_2,\ldots,n_k}$, where $k$, $n_1$, $n_2$, $\ldots,n_k$ are positive integers satisfying $N = \sum_{i=1}^k n_i$.

%Furthermore, since there might have repeated values for $D_x$, we assume there are $N'$ unique values of $D_x$ and $N' \leq N$. Note that the allowable discretization edges happens between two consective values of non-repeated value of $D_x$. For example, if $D_X = \{ 1.0,1.0,2.0\}$, then the only possible discretization edge is on $1.5$.

@@

The global minimum of Equation~\ref{eqn:MDL} can be found using dynamic programming.
For a given Bayesian network, the first three terms only depend on the number of discretization intervals $|X|$ and the fourth term is cumulative over the intervals.
Hence, if ${\Lambda = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_k}}$ is an MDL optimal discretization policy with $k$ intervals, then ${\Lambda' = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_{k-1}}}$ is an MDL optimal discretization policy for the corresponding subproblem with $k-1$ intervals.
This procedure takes runtime $\bigo \paren{ n^3 + \paren{ n_c \cdot {L}^{n_s}  + {L}^{n_p}} \cdot n^2 }$.
All variables are defined in Section \ref{subsec:algo}.
Therefore, the runtime of the proposed method in this paper is less than the MDL discretization method by $\bigo \paren{n^3}$, since the former has a more efficient form of dynamical programming.
For faster methods of MDL discretization with suboptimal results also, see \citet{Friedman_1996}.

@@

\begin{minipage}{\textwidth}
  \begin{minipage}[b]{0.49\textwidth}
    \centering
    \begin{tabular}{cl}
      \toprule
      Variable               & Distribution                                   \\ \midrule
      \multirow{2}{*}{$A$}&$P(a^1) = 0.5$     \\
      & $P( a^2) = 0.5$     \\ \midrule

      \multirow{2}{*}{$B$} & $p(b \mid A = a^1) \sim U[0,1]$\\
      & $p(b \mid A = a^2) \sim U[1,2]$\\
      \bottomrule
    \end{tabular}
    \vspace{1cm}
    \captionof{table}{Probability distributions for $A \rightarrow B$, in which $A$ is discrete and $B$ is continuous. The marginal distribution for $B$ is $U[0,2]$.}
    \label{table:marginal_counter}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\textwidth}
    \centering
    \vspace{1em}
    \resizebox {\columnwidth}{!}{
      \begin{tikzpicture}
      \begin{axis}[xlabel=$b$, ylabel=probability density, ymin=0, ymax=1.5,
        legend style={
          draw=none,
          at={(0.95,0.85)},
          anchor=east,
          legend columns=1,
          font=\small,
          /tikz/every even column/.append style={column sep=0.5cm},
          /pgfplots/legend image code/.code={%
            \draw[##1,/tikz/.cd,yshift=-0.25em]
            (0cm,0cm) rectangle (0.8em,0.8em);
          },
        },
      ]
      \addplot+[ybar interval, mark=none, black!80, fill=black!80] plot coordinates
        {(0,1) (1,1)};
        \addplot+[ybar interval, mark=none, black!40, fill=black!40] plot coordinates
        {(1,1) (2,1)};
        \legend{$P(b\mid a^1)$, $P(b\mid a^2)$}
      \end{axis}
      \end{tikzpicture}
    }
    \captionof{figure}{The resulting conditional probability distribution for $B$. Marginal methods see one uniform distribution.}
    \label{fig:marginal_counter}
  \end{minipage}
\end{minipage}

>>>>

\jairheading{59}{2017}{103-132}{10/16}{06/17}
\ShortHeadings{Learning Discrete Bayesian Networks from Continuous Data}
{Chen, Wheeler, \& Kochenderfer}
\firstpageno{103}

@@

\author{\name Yi-Chun Chen \email yichunc0419@alumni.stanford.edu \\
	\addr Institute of Computational and Mathematical Engineering,\\
	Stanford University, Stanford, CA  94035 USA
	\AND
	\name Tim A. Wheeler \email wheelert@stanford.edu \\
	\addr Department of Aeronautics and Astronautics,\\ Stanford University,
	Stanford, CA  94035 USA
	\AND
	\name Mykel J. Kochenderfer \email mykel@stanford.edu \\
	\addr Department of Aeronautics and Astronautics,\\ Stanford University,
	Stanford, CA  94035 USA}
%\authorrunning{Short form of author list} % if too long for running head

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

The global minimum of Equation~\ref{eqn:MDL} can be found using dynamic programming.
For a given Bayesian network, the first three terms only depend on the number of discretization intervals $|X|$ and the fourth term is cumulative over the intervals.
Hence, if ${\Lambda = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_k}}$ is an MDL optimal discretization policy with $k$ intervals, then ${\Lambda' = \anglebrackets{\gamma_1, \gamma_2, \ldots, \gamma_{k-1}}}$ is an MDL optimal discretization policy for the corresponding subproblem with $k-1$ intervals.
This procedure takes runtime $\bigo \paren{ n^3 + \paren{ n_c \cdot {L}^{n_s}  + {L}^{n_p}} \cdot n^2 }$.
All variables are defined in Section \ref{subsec:algo}.
Therefore, the runtime of the proposed method in this paper is less than the MDL discretization method by $\bigo \paren{n^3}$, since the former has a more efficient form of dynamical programming.
There also exist faster methods of MDL discretization, but they lead to suboptimal results \citep{Friedman_1996}.

@@

\begin{table}
	\centering
	\begin{tabular}{cl}
		\toprule
		Variable               & Distribution                                   \\ \midrule
		\multirow{2}{*}{$A$}&$P(a^1) = 0.5$     \\
		& $P( a^2) = 0.5$     \\ \midrule

		\multirow{2}{*}{$B$} & $p(b \mid A = a^1) \sim U[0,1]$\\
		& $p(b \mid A = a^2) \sim U[1,2]$\\
		\bottomrule
	\end{tabular}
	\vspace{1cm}
	\caption{Probability distributions for $A \rightarrow B$, in which $A$ is discrete and $B$ is continuous. The marginal distribution for $B$ is $U[0,2]$.}
\label{table:marginal_counter}
\end{table}

\begin{figure}[ht]
    \centering
    \vspace{1em}
%    \resizebox {\columnwidth}{!}{
    	\begin{tikzpicture}
    	\begin{axis}[xlabel=$b$, ylabel=probability density, ymin=0, ymax=1.5,
    	legend style={
    		draw=none,
    		at={(0.95,0.85)},
    		anchor=east,
    		legend columns=1,
    		font=\small,
    		/tikz/every even column/.append style={column sep=0.5cm},
    		/pgfplots/legend image code/.code={%
    			\draw[##1,/tikz/.cd,yshift=-0.25em]
    			(0cm,0cm) rectangle (0.8em,0.8em);
    		},
    	},
    	]
    	\addplot+[ybar interval, mark=none, black!80, fill=black!80] plot coordinates
    	{(0,1) (1,1)};
    	\addplot+[ybar interval, mark=none, black!40, fill=black!40] plot coordinates
    	{(1,1) (2,1)};
    	\legend{$P(b\mid a^1)$, $P(b\mid a^2)$}
    	\end{axis}
    	\end{tikzpicture}
%    }
    \caption{The resulting conditional probability distribution for $B$. Marginal methods see one uniform distribution.}
    \label{fig:marginal_counter}
\end{figure}

%
%\begin{minipage}{\textwidth}
%  \begin{minipage}[b]{0.49\textwidth}
%    \centering
%    \begin{tabular}{cl}
%      \toprule
%      Variable               & Distribution                                   \\ \midrule
%      \multirow{2}{*}{$A$}&$P(a^1) = 0.5$     \\
%      & $P( a^2) = 0.5$     \\ \midrule
%
%      \multirow{2}{*}{$B$} & $p(b \mid A = a^1) \sim U[0,1]$\\
%      & $p(b \mid A = a^2) \sim U[1,2]$\\
%      \bottomrule
%    \end{tabular}
%    \vspace{1cm}
%    \captionof{table}{Probability distributions for $A \rightarrow B$, in which $A$ is discrete and $B$ is continuous. The marginal distribution for $B$ is $U[0,2]$.}
%    \label{table:marginal_counter}
%  \end{minipage}
%  \hfill
%  \begin{minipage}[b]{0.49\textwidth}
%    \centering
%    \vspace{1em}
%    \resizebox {\columnwidth}{!}{
%      \begin{tikzpicture}
%      \begin{axis}[xlabel=$b$, ylabel=probability density, ymin=0, ymax=1.5,
%        legend style={
%          draw=none,
%          at={(0.95,0.85)},
%          anchor=east,
%          legend columns=1,
%          font=\small,
%          /tikz/every even column/.append style={column sep=0.5cm},
%          /pgfplots/legend image code/.code={%
%            \draw[##1,/tikz/.cd,yshift=-0.25em]
%            (0cm,0cm) rectangle (0.8em,0.8em);
%          },
%        },
%      ]
%      \addplot+[ybar interval, mark=none, black!80, fill=black!80] plot coordinates
%        {(0,1) (1,1)};
%        \addplot+[ybar interval, mark=none, black!40, fill=black!40] plot coordinates
%        {(1,1) (2,1)};
%        \legend{$P(b\mid a^1)$, $P(b\mid a^2)$}
%      \end{axis}
%      \end{tikzpicture}
%    }
%    \captionof{figure}{The resulting conditional probability distribution for $B$. Marginal methods see one uniform distribution.}
%    \label{fig:marginal_counter}
%  \end{minipage}
%\end{minipage}
